---
author: "Glenn Morton, Yeji Bae"
date: "last-modified"
output:
  html_document: default
always_allow_html: true
---
# Linux/Unix Command Line Cheat Sheet {#sec-linuxcheatsheet}

```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE, include=TRUE)
library(knitr)
library(kableExtra)
```

Adapted from GettingGeneticsDone.blogspot.com

```{r, echo=FALSE, eval=TRUE}
filePath <- "tables/LinuxCheatSheet_commands.txt"
t1 <- read.table(file=filePath, sep="\t", header = TRUE, check.names = FALSE, stringsAsFactors = FALSE)
footnote <- tail(t1, n = 1)[1]
t1 <- head(t1, n=-1)
kable(t1, align = 'l') %>%
  add_footnote(footnote, notation = 'none') %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE, html_font = "\"Trebuchet MS\", verdana, sans-serif")
```

#### Absolute vs relative paths.

Let’s say you are here: `/home/turnersd/scripts/`. If you wanted to go to `/home/turnersd/`, you could type: `cd /home/turnersd/.` Or you could
use a relative path. `cd .. `(two periods) will take you one directory “up” to the parent directory of the current directory.

* . (a single period) means the current directory

* .. (two periods) means the parent directory

* ~ means your home directory

```{r, echo=FALSE, eval=TRUE}
filePath <- "tables/LinuxCheatSheet_absoluteVsRelative.txt"
t1 <- read.table(file=filePath, sep="\t", header = FALSE, check.names = FALSE, stringsAsFactors = FALSE)
title = c(2)
names(title) <- head(t1, n = 1)[1]
t1 <- tail(t1, n = nrow(t1)-1)
kable(t1, align = 'l', col.names = NULL, row.names = FALSE) %>%
  add_header_above(header = title) %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE, html_font = "\"Trebuchet MS\", verdana, sans-serif")
```

```{r, echo=FALSE, eval=TRUE}
filePath <- "tables/LinuxCheatSheet_wildcards.txt"
t1 <- read.table(file=filePath, sep="\t", header = FALSE, check.names = FALSE, stringsAsFactors = FALSE)
title = c(2)
names(title) <- head(t1, n = 1)[1]
t1 <- tail(t1, n = nrow(t1)-1)
kable(t1, align = 'l', col.names = NULL, row.names = FALSE) %>% 
  add_header_above(header = title) %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE, html_font = "\"Trebuchet MS\", verdana, sans-serif")
```